Cycles of the logistic map

### 2013-05-27 doc/pc-ijbc.tex
pc-ijbc: remove figure and tables

\newcommand{\vx}{\vct x}
\newcommand{\vxn}[1]{\vx_{#1}}
\newcommand{\vyn}[1]{\vct y_{#1}}
\newcommand{\Tr}{\mathrm{Tr}}
% basis set of cyclic polynomials
\newcommand{\Pset}{\mathcal P}
\newcommand{\NB}{N_\Pset}
\newcommand{\LB}{L_\Pset}
\newcommand{\lam}{(\lambda)}
\newcommand{\Rlam}{(R, \lambda)}
\newcommand{\rlam}{(r, \lambda)}

@@

Consider the logistic map \cite{may, strogatz}:
%
\begin{equation}
  x_{k+1} = f(x_k) \equiv r \, x_k \, ( 1 - x_k ).
\label{eq:logmap}
\end{equation}
%
The iterated sequence $x_1$,
  $x_2 = f(x_1)$,
  $x_3 = f(x_2)$,
  $\ldots$,
can be conveniently visualized on the cobweb plot,
  as shown in \reffig{cobweb}.
%
Starting from $(x_1, x_1)$ on the diagonal,
  each vertical arrow takes $(x_k, x_k)$ to $(x_k, y)$,
  where $y = x_{k+1} = f(x_k)$;
by the following horizontal arrow,
  it returns to the diagonal at
  $(y, y) = (x_{k+1}, x_{k+1})$,
  which starts the next iteration.
%
%
\begin{figure}[h]
  \begin{center}
  \begin{minipage}{\linewidth}
        \includegraphics[angle=-90, width=\linewidth]{cobweb.pdf}
  \end{minipage}%
  \end{center}
  \caption{\label{fig:cobweb}
  Cobweb plot of the logistic map.}
\end{figure}
%
%
After many iterations, there can be three ultimate patterns:
  (i) a \emph{fixed point} (including infinity),
      e.g., \reffig{cobweb}(a);
  (ii) a periodic \emph{cycle},
      e.g., \reffigs{cobweb}(b)-(e);
or
  (iii) a chaotic trajectory,
      e.g., \reffig{cobweb}(f).
We will focus on the first two cases here.

A fixed point is a solution of $x^* = f(x^*)$.
%
It is stable if a small deviation from $x^*$
  gets reduced by successive iterations.
%
For a differentiable $f$,
  this is equivalent to $|f'(x^*)| \le 1$
  \cite{strogatz}.

An $n$-cycle is a sequence $x_1, \dots, x_n$,
  such that $n$ is the smallest positive integer
  that satisfies $x_1 = x_{n+1} = f^n(x_1)$,
  where $f^n$ is the $n$th iterate of $f$,
  e.g., $f^3(x) = f(f(f(x)))$.
Obviously, any $x_k$ in an $n$-cycle of $f$ is a fixed point of $f^n$
%
\big[but the converse is untrue, for a fixed point of $f^n$
  can also be a fixed point of $f^d$ for a divisor $d$ of $n$:
  if $f^d(x) = x$, then $f^n(x) = f^d(\cdots f^d(x)\cdots) = x$\big].
%
%
Thus, a cycle can be classified as stable or unstable
  according to the corresponding fixed point of $f^n$:
  a stable cycle requires
  $\big| \frac {d} {dx} f^n(x_1) \big| \le 1$,
  or by the chain rule,
%
%
%
\begin{equation}
  \Big| f'(x_1) \cdots f'(x_n) \Big| \le 1,
\label{eq:der}
\end{equation}
%
%
The onset and bifurcation points
  are defined at the loci
  where $\frac {d} {dx} f^n(x_1)$ reaches $+1$ and $-1$,
  respectively \cite{strogatz}.
%
%
%
The cycle and its stability depend on the parameter $r$.
%
The problem of identifying
  all regions of $r$ that allow stable $n$-cycles,
  has been the objective of several previous studies
  \cite{brown1, brown2, stephenson1, stephenson2, stephenson3,
  saha, bechhoefer, gordon, burm, zhang, bailey1, bailey2, kk1, lewis}.
Below we present a simple algorithm to do so.

@@

We will illustrate the algorithm on the logistic map % \cite{may, strogatz}
  \refeq{logmap}.
%
The generalizations are straightforward.
%
If $r$ is real,
  we will find windows $(r_a, r_b)$,
    within which stable $n$-cycles can exist.
%
%The window boundaries can be classified
%  to onset and bifurcation points according to \refeq{der};
Here, if $r > 0$, then
  $r_a$ ($r_b$) are the onset (bifurcation) points.
%
There are generally multiple such windows for a single $n$,
  but all onset points satisfy the same polynomial equation,
  and all bifurcation points another.
%
Our goal is thus to find the two polynomials for a given $n$.

To simplify the calculation, we first change variables \cite{mandelbrot, brown1, brown2, saha} as
%
%\begin{equation}
\[
    x^{\mathrm{new}} \equiv r(x^{\mathrm{old}} - 1/2),
    \quad R \equiv r(r-2)/4,
\]
%\label{eq:chvars}
%\end{equation}
%
and rewrite the map, in terms of $x^{\mathrm{new}}$, as
%
\begin{equation}
  x_{k+1} = f(x_k) \equiv R - {x_k}^2.
  %\tag{$\ref{eq:logmap}'$}
\label{eq:logmaps}
\end{equation}
%
%\refeq{logmaps} simplifies \refeq{logmap} as $x_k$ now only occurs once.
%
We will solve the cycle boundaries as zeros of some polynomials of $R$,
and the corresponding polynomials for $r$ can be obtained by $R \rightarrow r(r-2)/4$.
%
%the numerical $r$ values can always be obtained from $r = 1 + \sqrt{1 + 4R}$.
%

\subsection{Overall plan}

Since the $n$-cycles points constitute a subset of the fixed points of $f^n$,
we will first find
  the polynomials at the stability boundaries
  for the fixed points of $f^n$
  (\refsecs{cyclic}-\ref{sec:examples}).
The contributions from shorter $d$-cycles ($d|n$)
  are lower-degree factors of the polynomials,
  and can be readily removed (\refsec{primfac}).

At the first glance, the polynomials of the fixed points of $f^n$
  can be obtained by brute force.
We can solve
  \refeqs{logmaps} and express $x_1, \ldots, x_n$
  in terms of $R$,
  and then plug the solution into \req{der}.
The result would contain $R$ only,
  and is therefore the answer.
But since \refeqs{logmaps} are nonlinear,
  it is nontrivial to reduce the final equation of $R$
  into a polynomial one.
Nonetheless, on a computer, one can
  construct a polynomial Gr\"obner basis \cite{kk1},
  which gradually eliminates $x_k$ out of the $n+1$ equations.
If $x_2, \dots, x_n$ are eliminated first,
  there are only two equations left:
  $f^n(x_1) = x_1$
  and
  $\frac{d}{dx}f^n (x_1) = \pm 1$
  [\refeq{der}];
  the final elimination of $x_1$ can then be accomplished
  by setting the resultant of the two equations
  to zero \cite{burm},
  leading to the desired polynomial equation of $R$.
%
These approaches, however, do not
  exploit the cyclic structure of \refeqs{logmaps},
  and can thus be improved by the following alternative.

Instead of solving \refeqs{logmaps} for $x_k$,
  we will derive a set of equations
  in terms of the cyclic polynomials $\{C_p\}$ of $x_k$
  (an example of a cyclic polynomial is $x_1 x_2 + x_2 x_3 + \dots + x_n x_1$).
%
The advantage is that the new set of equations are
  \emph{homogeneous}, \emph{linear} and \emph{complete},
  i.e., there is an equation
  $\sum_q A_{pq} C_q = 0$ for every cyclic polynomial labeled by $p$.
%
Now the matrix formed by the coefficients $A_{pq}$
  must have a zero determinant,
  for the cyclic polynomials $\{C_p\}$ are not zeros altogether.
%
Thus, the zero-determinant condition $|\mathbf A| = 0$ gives
  the required polynomial equation of $R$.
%whose roots contain all fixed points of $f^n$.
%

@@

A cyclic polynomial is an invariant
  under the cycling of variables
  $x_1 \rightarrow x_2, x_2 \rightarrow x_3,
  \ldots, x_n \rightarrow x_1$,
  e.g., $a = x_1 x_2 + x_2 x_3 + x_3 x_4 + x_4 x_1$
  and $b = x_1 x_3 + x_2 x_4$ for $n = 4$.
%
It is, however, unnecessarily symmetric
  (invariant under the exchange of any two $x_k$ and $x_{j}$),
  e.g., $a$ and $b$ are not symmetric,
  although $a + b$ is.

A cyclic polynomial can be generated by summing over
  distinct cyclic versions of a simpler polynomial of $x_k$, or a \emph{generator}.
The generators of the above $a$ and $b$
  are $x_1 x_2$ and $x_1 x_3$, respectively.
%For example,
%$x_1 x_2$ is a generator of $a$;
%$x_1 x_3$ is that of $b$;
%and
%$x_1 x_2 + \frac{1}{2} x_1 x_3$ is that of $a + b$.
%Note that, according this definition, the coefficient before $x_1 x_3$
%  is 1 in the second case
%  for there are only two distinct versions,
%  but $\frac{1}{2}$ in the third case
%  for there are four. % distinct versions of the generator.
%
%
%
Cyclic polynomials
  generated from monomials of unit coefficient,
  such as $a$ and $b$, % but not $a + b$ in the above example)
  can be labeled as follows.
If we write the monomial generator as
  ${x_1}^{e_1} {x_2}^{e_2} \dots {x_n}^{e_n}$,
  then the sequence $p$ of indices with
  $e_1$ 1's, $e_2$ 2's, \ldots, $e_n$ $n$'s
  serves as the label of the cyclic polynomial $C_p(\vxn{n})$,
  where $\vxn{n} \equiv \{x_1, \ldots, x_n\}$
For example,
$C_{12}(\vxn{n})  = x_1 x_2 + x_2 x_3 + \dots + x_n x_1$
and
$C_{112}(\vxn{n}) = {x_1}^2 x_2 + {x_2}^2 x_3 + \dots + {x_n}^2 x_1$.
%
Below, we may drop the subscript ``$_n$'' or
  even the argument ``$(\vxn{n})$'' entirely when it is clear.
In case the cyclic polynomial has several generators,
  we pick the one with the smallest $p$
  in the sense of lexicographic order,
  e.g., we choose $x_1 x_2$ over $x_2 x_3$ for $C_{12}$
  because $12 < 23$ ($n \ge 3$).
Finally, we add $C_0 \equiv 1$ for completeness.
%
%
%

@@

A cyclic polynomial is \emph{square-free}
  if it has no square or higher powers of $x_k$,
e.g., $C_{12} = x_1 x_2 + \cdots$ is square-free,
  $C_{112} = {x_1}^2 x_2 + \cdots$ is not.
%see \reftab{sqrfreepoly} for more examples.
The label $p$ of a square-free polynomial
  has no repeated index.
We denote the set of all square-free $p$ by
$\Pset = \{0, 1, 12, 13, \ldots, 123, 124, \ldots, 12\dots n\}$,
and its size by $\NB$.
% is the number of square-free cyclic polynomials.
%
%
The set of square-free cyclic polynomials
  furnishes a basis for expanding cyclic polynomials:

\begin{theorem}
  For the logistic map \refeq{logmaps},
  any cyclic polynomial $K(\vxn{n})$
  formed by the $n$-cycle points
  $\vxn{n} = \{x_1, \ldots, x_n\}$
  is a linear combination of
  the square-free cyclic polynomials $C_p(\vxn{n})$:
\[
  K(\vxn{n}) = \sum_{p \in \Pset} f_p(R) \, C_p(\vxn{n}),
\]
  where $\Pset$ % = \{0, 1, 12, 13, \ldots, 12\dots n\}$ is
  is the set of indices of all square-free cyclic polynomials,
  and $f_p(R)$ are polynomials of $R$.
  \label{thm:sqrfree}
\end{theorem}
%
%
\begin{proof}
Consider the following square-free reduction.
%
Given a cyclic polynomial $K(\vxn{n})$,
  we recursively apply \refeq{logmaps} as
  $x_k^2 \rightarrow R - x_{k+1}$,
  until all squares or higher powers of $x_k$ are eliminated.
The process will not last indefinitely
  for each substitution reduces the degree in $x_k$ %(no matter which $k$)
  by one.
Since both the replacement rules and the original polynomial are cyclic,
  so is the final square-free polynomial.
Terms involving no $x_k$
  are collected to serve as the coefficient before $C_0(\vxn{n}) \equiv 1$.
The coefficients $f_p$ are polynomials of $R$,
  for $R$ is the only new variable introduced by the substitutions.
\end{proof}
%
%
For example, the cyclic polynomial
  $K(\vxn{2}) = {x_1}^2 \, x_2 + {x_2}^2 \, x_1$
can be written as $K = (R + 1) \, C_{1} - 2 R \, C_0$
for
${x_1}^2 \, x_2 = R \, x_2 - {x_2}^2 = R \, x_2 - R + x_1$ and
${x_2}^2 \, x_1 = R \, x_1 - R + x_2$.

\refthm{sqrfree}
  shows that any cyclic polynomial can be expanded
  as a combination of the square-free ones.
Below we show that at the onset and bifurcation points,
  the square-free cyclic polynomials $C_p$ are themselves
  linearly connected by $\NB$ homogeneous equations.
The matrix of the connecting coefficients thus has a vanishing determinant.
This condition furnishes the desired polynomial equation.

@@

First, the derivative of $f^n$ is a cyclic polynomial:
%
%
\begin{equation}
  \Lambda(\vxn{n})
   = f'(x_1) \dots f'(x_n)
   = (-2)^n x_1 \dots x_n.
\label{eq:logder}
\end{equation}
%
%
For any square-free cyclic polynomial $C_p(\vxn{n})$,
the product $\Lambda(\vxn{n}) \, C_p(\vxn{n})$ is also cyclic,
  and can thus be expanded by \refthm{sqrfree} as
\begin{equation}
  \Lambda(\vxn{n}) \, C_p(\vxn{n}) = \sum_{q \in \Pset} T_{pq}(R) \, C_q(\vxn{n}),
\label{eq:xcp0}
\end{equation}
where $T_{pq}(R)$ is a polynomial of $R$,
and $p, q \in \Pset$.

By \refeq{der},
  $\Lambda(\vxn{n})$ equals to $\lambda = +1$ or $-1$
  at the onset or bifurcation point,
  respectively;
  so \refeq{xcp0}
  becomes a homogeneous linear equation of
  $C_p(\vxn{n})$:
\begin{equation}
  \lambda \, C_p(\vxn{n}) = \sum_{q \in \Pset} T_{pq}(R) \, C_q(\vxn{n}),
\label{eq:xcp}
\end{equation}
%
%
%where $\lambda$ is $+1$ and $-1$ at the onset and bifurcation points, respectively;
or in matrix form,
\begin{equation}
  \big[ \lambda \, \vct I - \vct T(R) \big] \, \vct C = 0,
\tag{\ref{eq:xcp}$'$}
\end{equation}
%
%
where
$\vct I$ is the $\NB \times \NB$ identity matrix,
$\vct T(R) = \{T_{pq}(R)\}$ is an $\NB \times \NB$ matrix,
and
$\vct C = \{C_p(\vxn{n})\}$ is an $\NB$-dimensional column vector.

Since a set of homogeneous linear equations
  has a nontrivial solution
  ($C_p \ne 0$ for at least one $p$)
  only if the determinant of the coefficient matrix
  is zero, we have
\begin{equation}
  A_n\Rlam \equiv \big| \lambda \, \vct I - \vct T(R) \big| = 0.
\label{eq:xr}
\end{equation}
%
Now $A_n\Rlam$, from expanding the determinant,
  is a polynomial of $R$ and $\lambda$,
  the equation is the desired answer to our problem.
%
%The attached subscript $n$ allows us
%  to use the determinant later with
%  similar determinants of divisors $d$ of $n$.
%
To summarize, we have
\begin{theorem}
  At the onset and bifurcation points,
  the square-free cyclic polynomials $C_p(\vxn{n})$
  of the cycle points $\vxn{n} = \{x_k\}$
  are linearly related as
  $\lambda \, C_p(\vxn{n}) = \sum_{q \in \Pset} T_{pq}(R) \, C_q(\vxn{n})$,
  with $\lambda$ being $+1$ and $-1$, respectively,
  and $T_{pq}(R)$ being the coefficients
  from the square-free reduction of the cyclic polynomial
  $C_p(\vxn{n}) \, \prod_k f'(x_k)$. %[\refeq{logder}].
  Thus, the fixed points of $f^n$
  are marginally stable when $R$ satisfies
  the characteristic equation
  $\big| \lambda \, \vct I - \vct T(R) \big| = 0$.
  \label{thm:main}
\end{theorem}

%\begin{remark}
More generally, the algorithm can trace a contour of marginal stability of $R$
  in the complex plane
  for complex cycles of $\vxn{n}$
  if $\lambda$ is set to $\exp(i\theta)$ with $\theta \in [0, 2\pi]$.
  The region enclosed by the contour corresponds to
    a bulb in the inverted Mandelbrot set
  (the inversion $z_k = -x_k, c = -R$
  reduces \refeq{logmaps} to the standard Mandelbrot form: $z_{k+1} = c + {z_k}^2$
  \cite{mandelbrot}),
  see ref. \cite{stephenson2, stephenson3}. % and \reffig{logbulb}.
%\end{remark}

@@

We illustrate the above algorithm on several small-$n$ cases.
%
It is still helpful to have a mathematical software
  at hand to verify certain steps
  (e.g., in computing the determinants and their factorization).

For $n = 1$ (the fixed point), we have
  $C_0 = 1$, $\Lambda = -2 \, C_1 = -2 \, x_1$, and
\[
  \Lambda
  \left( \begin{array}{c}
  C_0 \\
  C_1
  \end{array} \right)
  =
  \left( \begin{array}{cc}
  0     & -2 \\
  -2R   &  2
  \end{array}\right)
  \left( \begin{array}{c}
  C_0 \\
  C_1
  \end{array} \right).
\]
With $\Lambda \rightarrow \lambda$,
  the characteristic equation (\ref{eq:xr}) reads
\begin{equation}
A_1\Rlam
  %=\left| \begin{array}{cc}
  %\lambda     & 2          \\
  %2 R         & \lambda - 2
  %\end{array} \right|
  = -4R - 2\lambda + \lambda^2
  = 0.
  \tag{\ref{eq:xr}-1}
\label{eq:xr1}
\end{equation}
So $R = (\lambda^2 -2\lambda)/4$,
  and the fixed point
  %begins at $R_a = -1/4$ ($\lambda = +1$), and
  becomes unstable at $R_b = 3/4$ where $\lambda = -1$.

>>>>

\newcommand{\vx}[1]{\vct x_{#1}}
\newcommand{\vy}[1]{\vct y_{#1}}
\newcommand{\vz}[1]{\vct z_{#1}}
\newcommand{\Tr}{\mathrm{Tr}}
% basis set of cyclic polynomials
\newcommand{\Pset}{\mathcal P}
\newcommand{\NB}{N_\Pset}
\newcommand{\LB}{L_\Pset}
\newcommand{\lam}{(\lambda)}
\newcommand{\Rlam}{(R, \lambda)}
\newcommand{\rlam}{(r, \lambda)}

@@

Consider the logistic map \cite{may, strogatz}:
%
\begin{equation}
  x_{k+1} = f(x_k) \equiv r \, x_k \, ( 1 - x_k ).
\label{eq:logmap}
\end{equation}
%
%The iterated sequence $x_1$,
%  $x_2 = f(x_1)$,
%  $x_3 = f(x_2)$,
%  $\ldots$,
%can be conveniently visualized on the cobweb plot,
%  as shown in \reffig{cobweb}.
%%
%Starting from $(x_1, x_1)$ on the diagonal,
%  each vertical arrow takes $(x_k, x_k)$ to $(x_k, y)$,
%  where $y = x_{k+1} = f(x_k)$;
%by the following horizontal arrow,
%  it returns to the diagonal at
%  $(y, y) = (x_{k+1}, x_{k+1})$,
%  which starts the next iteration.
%
%
%\begin{figure}[h]
%  \begin{center}
%  \begin{minipage}{\linewidth}
%        \includegraphics[angle=-90, width=\linewidth]{cobweb.pdf}
%  \end{minipage}%
%  \end{center}
%  \caption{\label{fig:cobweb}
%  Cobweb plot of the logistic map.}
%\end{figure}
%
%
After many iterations, there can be three ultimate patterns:
  (i) a \emph{fixed point} (including infinity),
      %e.g., \reffig{cobweb}(a);
  (ii) a periodic \emph{cycle},
      %e.g., \reffigs{cobweb}(b)-(e);
or
  (iii) a chaotic trajectory.
      %e.g., \reffig{cobweb}(f).
We only discuss the first two cases here.

A fixed point is a solution of $x^* = f(x^*)$.
%
It is stable if a small deviation from $x^*$
  gets reduced by successive iterations.
%
For a differentiable $f$,
  this is equivalent to $|f'(x^*)| \le 1$
  \cite{strogatz}.
%
An $n$-cycle is a sequence $x_1, \dots, x_n$,
  such that $n$ is the smallest positive integer
  that satisfies $x_1 = x_{n+1} = f^n(x_1)$,
  where $f^n$ is the $n$th iterate of $f$,
  e.g., $f^3(x) = f(f(f(x)))$.
Obviously, any $x_k$ in an $n$-cycle of $f$ is a fixed point of $f^n$.
%
%\big[but the converse is untrue, for a fixed point of $f^n$
%  can also be a fixed point of $f^d$ for a divisor $d$ of $n$:
%  if $f^d(x) = x$, then $f^n(x) = f^d(\cdots f^d(x)\cdots) = x$\big].
%
%
Thus, a cycle can be classified as stable or unstable
  according to the corresponding fixed point of $f^n$:
  a stable cycle requires
  $\big| \frac {d} {dx} f^n(x_1) \big| \le 1$,
  or by the chain rule,
%
%
%
\begin{equation}
  \Big| f'(x_1) \cdots f'(x_n) \Big| \le 1,
\label{eq:der}
\end{equation}
%
%
The onset and bifurcation points
  are defined at the loci
  where $\frac {d} {dx} f^n(x_1)$ reaches $+1$ and $-1$,
  respectively \cite{strogatz}.
%
%
%
The cycle and its stability depend on the parameter $r$,
%
and a stable cycle corresponds to a window
  between the onset point $r_a$ and bifurcation point $r_b$.
%
There are generally multiple such windows for a single $n$,
  but all $r_a$ are zeros of a polynomial,
  and all $r_b$ another.
%
Our goal is thus to find the two polynomials for a given $n$.
%
The problem has been tackled by several previous studies
  \cite{brown1, brown2, stephenson1, stephenson2, stephenson3,
  saha, bechhoefer, gordon, burm, zhang, bailey1, bailey2, kk1, lewis}.
Below we present a simple algorithm to do so.
%
%
%
%The window boundaries can be classified
%  to onset and bifurcation points according to \refeq{der};
%Here, if $r > 0$, then
%  $r_a$ ($r_b$) are the onset (bifurcation) points.
%

@@

To simplify the calculation, we first change variables \cite{mandelbrot, brown1} as
%
%\begin{equation}
\[
    x^{\mathrm{new}} \equiv r(x^{\mathrm{old}} - 1/2),
    \quad R \equiv r(r-2)/4,
\]
%\label{eq:chvars}
%\end{equation}
%
and rewrite the map, in terms of $x^{\mathrm{new}}$, as
%
\begin{equation}
  x_{k+1} = f(x_k) \equiv R - {x_k}^2.
  %\tag{$\ref{eq:logmap}'$}
\label{eq:logmaps}
\end{equation}
%
%\refeq{logmaps} simplifies \refeq{logmap} as $x_k$ now only occurs once.
%
We will solve the cycle boundaries as zeros of some polynomials of $R$,
and the corresponding polynomials for $r$ can be obtained by $R \rightarrow r(r-2)/4$.
%
%the numerical $r$ values can always be obtained from $r = 1 + \sqrt{1 + 4R}$.
%

\subsection{Strategy}

Since the $n$-cycles points constitute a subset of the fixed points of $f^n$,
we will first find
  the polynomials at the stability boundaries
  for the fixed points of $f^n$
  (\refsecs{cyclic}-\ref{sec:examples}).
The polynomials contain
  factors, contributed by short $d$-cycles with $d|n$
  (which are also fixed points of $f^n$).
The factors are removed in \refsec{primfac}.

At the first glance, the polynomials of the fixed points of $f^n$
  can be obtained by brute force.
We can solve
  \refeqs{logmaps} and express $x_1, \ldots, x_n$
  in terms of $R$,
  and then plug the solution into \refeq{der}.
The result would contain $R$ only,
  and is therefore the answer.
But since \refeqs{logmaps} are nonlinear,
  it is nontrivial to reduce the final equation of $R$
  into a polynomial one.
Nonetheless, on a computer, one can
  construct a polynomial Gr\"obner basis \cite{kk1},
  which gradually eliminates $x_k$ out of the $n+1$ equations.
If $x_2, \dots, x_n$ are eliminated first,
  there are only two equations left:
  $f^n(x_1) = x_1$
  and
  $\frac{d}{dx}f^n (x_1) = \pm 1$
  [\refeq{der}];
  the final elimination of $x_1$ can then be accomplished
  by setting the resultant of the two equations
  to zero \cite{burm},
  leading to the desired polynomial equation of $R$.
%
These approaches, however, do not
  exploit the cyclic structure of \refeqs{logmaps},
  and can thus be improved by the following alternative.

Instead of solving \refeqs{logmaps} for $x_k$,
  we will derive a set of equations
  in terms of the cyclic polynomials $\{C_p\}$ of $x_k$
  (an example of a cyclic polynomial is
  $x_1 x_2 + x_2 x_3 + \dots + x_n x_1$
  with $n$ terms).
%
The advantage is that the new set of equations are
  \emph{homogeneous}, \emph{linear} and \emph{complete},
  i.e., there is an equation
  $\sum_q A_{pq} C_q = 0$ for every cyclic polynomial labeled by $p$.
%
Now the matrix formed by the coefficients $A_{pq}$
  must have a zero determinant,
  for the cyclic polynomials $\{C_p\}$ are not zeros altogether.
%
Thus, the zero-determinant condition $|\mathbf A| = 0$ gives
  the required polynomial equation of $R$.
%whose roots contain all fixed points of $f^n$.
%

@@

A cyclic polynomial is an invariant
  under the cycling of variables
  $x_1 \rightarrow x_2, x_2 \rightarrow x_3,
  \ldots, x_n \rightarrow x_1$,
  e.g., $a = x_1 x_2 + x_2 x_3 + x_3 x_4 + x_4 x_1$
  and $b = x_1 x_3 + x_2 x_4$ for $n = 4$.
%
It may not, however, be symmetric
  (invariant under the exchange of any two $x_i$ and $x_{j}$),
  e.g., $a$ and $b$ are not symmetric,
  only $a + b$ is.

A cyclic polynomial can be generated by summing over
  distinct cyclic versions of a simpler polynomial of $x_k$, or a \emph{generator}.
The generators of the above $a$ and $b$
  are $x_1 x_2$ and $x_1 x_3$, respectively.
%For example,
%$x_1 x_2$ is a generator of $a$;
%$x_1 x_3$ is that of $b$;
%and
%$x_1 x_2 + \frac{1}{2} x_1 x_3$ is that of $a + b$.
%Note that, according this definition, the coefficient before $x_1 x_3$
%  is 1 in the second case
%  for there are only two distinct versions,
%  but $\frac{1}{2}$ in the third case
%  for there are four. % distinct versions of the generator.
%
%
%
Cyclic polynomials
  generated from monomials of unit coefficient
  %such as $a$ and $b$, % but not $a + b$ in the above example)
  can be labeled as follows.
If we write the monomial generator as
  ${x_1}^{e_1} {x_2}^{e_2} \dots {x_n}^{e_n}$,
  then the sequence $p$ of indices with
  $e_1$ 1's, $e_2$ 2's, \ldots, $e_n$ $n$'s
  serves as the label of the cyclic polynomial $C_p(\vx{n})$,
  where $\vx{n} \equiv \{x_1, \ldots, x_n\}$.
%
For example,
$C_{12}(\vx{n})  = x_1 x_2 + x_2 x_3 + \dots + x_n x_1$
and
$C_{112}(\vx{n}) = {x_1}^2 x_2 + {x_2}^2 x_3 + \dots + {x_n}^2 x_1$.
%
Below, we may drop %the subscript ``$_n$'' or even the argument
  ``$(\vx{n})$'' when it is clear.
In case the cyclic polynomial has several generators,
  we pick the one with the smallest $p$
  in the sense of lexicographic order,
  e.g., we choose $x_1 x_2$ over $x_2 x_3$ for $C_{12}(\vx{n})$,
  because $12 < 23$.
%
Finally, we add $C_0(\vx{n}) \equiv 1$.
%
%
%

@@

A cyclic polynomial is \emph{square-free}
  if it has no square or higher powers of $x_k$,
e.g., $C_{12} = x_1 x_2 + \cdots$ is square-free,
  $C_{112} = {x_1}^2 x_2 + \cdots$ is not.
%see \reftab{sqrfreepoly} for more examples.
The label $p$ of a square-free polynomial
  has no repeated index.
We denote the set of all square-free $p$ by
$\Pset = \{0, 1, 12, 13, \ldots, 123, 124, \ldots, 12\cdots n\}$,
and its size by $\NB$.
% is the number of square-free cyclic polynomials.
%
%
The set of square-free cyclic polynomials
  furnishes a basis for expanding cyclic polynomials:

\begin{theorem}
  For the logistic map \refeq{logmaps},
  any cyclic polynomial $K(\vx{n})$
  formed by the $n$-cycle points
  $\vx{n} = \{x_1, \ldots, x_n\}$
  is a linear combination of
  the square-free cyclic polynomials $C_p(\vx{n})$:
\[
  K(\vx{n}) = \sum_{p \in \Pset} f_p(R) \, C_p(\vx{n}),
\]
  where $\Pset$ % = \{0, 1, 12, 13, \ldots, 12\dots n\}$ is
  is the set of indices of all square-free cyclic polynomials,
  and $f_p(R)$ are polynomials of $R$.
  \label{thm:sqrfree}
\end{theorem}
%
%
\begin{proof}
Consider the following square-free reduction.
%
Given a cyclic polynomial $K(\vx{n})$,
  we recursively apply \refeq{logmaps} as
  $x_k^2 \rightarrow R - x_{k+1}$,
  until all squares or higher powers of $x_k$ are eliminated.
The process will not last indefinitely
  for each substitution reduces the degree in $x_k$ %(no matter which $k$)
  by one.
Since both the replacement rules and the original polynomial are cyclic,
  so is the final square-free polynomial.
Terms involving no $x_k$
  are collected to serve as the coefficient before $C_0(\vx{n}) \equiv 1$.
The coefficients $f_p$ are polynomials of $R$,
  for $R$ is the only new variable introduced by the substitutions.
\end{proof}
%
%
For example, the cyclic polynomial
  $K(\vx{2}) = {x_1}^2 \, x_2 + {x_2}^2 \, x_1$
can be written as $K = (R + 1) \, C_{1} - 2 R \, C_0$
for
${x_1}^2 \, x_2 = R \, x_2 - {x_2}^2 = R \, x_2 - R + x_1$ and
${x_2}^2 \, x_1 = R \, x_1 - R + x_2$.

\refthm{sqrfree}
  shows that any cyclic polynomial can be expanded
  as a combination of the square-free ones.
At the onset and bifurcation points,
  the $\NB$ square-free cyclic polynomials $C_p$ are themselves
  linearly connected by \refeq{der}.
This condition yields a characteristic equation,
  which solves our problem.

@@

%The matrix of the connecting coefficients thus has a vanishing determinant.
%This condition furnishes the desired polynomial equation.

First, the derivative of $f^n$ is a cyclic polynomial:
%
%
\begin{equation}
  \Lambda(\vx{n})
   = f'(x_1) \dots f'(x_n)
   = (-2)^n x_1 \dots x_n.
\label{eq:logder}
\end{equation}
%
%
For any square-free cyclic polynomial $C_p(\vx{n})$,
the product $\Lambda(\vx{n}) \, C_p(\vx{n})$ is also cyclic,
  and can thus be expanded by \refthm{sqrfree} as
\begin{equation}
  \Lambda(\vx{n}) \, C_p(\vx{n}) = \sum_{q \in \Pset} T_{pq}(R) \, C_q(\vx{n}),
\label{eq:xcp0}
\end{equation}
where $T_{pq}(R)$ is a polynomial of $R$,
and $p, q \in \Pset$.

By \refeq{der},
  $\Lambda(\vx{n})$ equals to $\lambda = +1$ ($-1$)
  at the onset (bifurcation) point,
  respectively;
  so \refeq{xcp0}
  becomes a homogeneous linear equation of
  $C_p(\vx{n})$:
\begin{equation}
  \lambda \, C_p(\vx{n}) = \sum_{q \in \Pset} T_{pq}(R) \, C_q(\vx{n}),
\label{eq:xcp}
\end{equation}
%
%
%where $\lambda$ is $+1$ and $-1$ at the onset and bifurcation points, respectively;
or in matrix form,
\begin{equation}
  \big[ \vct T(R) - \lambda \, \vct I \big] \, \vct C = 0,
\tag{\ref{eq:xcp}$'$}
\end{equation}
%
%
where
$\vct I$ is the $\NB \times \NB$ identity matrix,
$\vct T(R) = \{T_{pq}(R)\}$ is an $\NB \times \NB$ matrix,
and
$\vct C = \{C_p(\vx{n})\}$ is an $\NB$-dimensional column vector.

Since a set of homogeneous linear equations
  has a nontrivial solution
  ($C_p \ne 0$ for at least one $p$)
  only if the determinant of the coefficient matrix
  is zero, we have
\begin{equation}
  A_n\Rlam \equiv \big| \vct T(R) - \lambda \, \vct I \big| = 0.
\label{eq:xr}
\end{equation}
%
This determinant $A_n\Rlam$
  is the desired polynomial of $R$ and $\lambda$.
%
%The attached subscript $n$ allows us
%  to use the determinant later with
%  similar determinants of divisors $d$ of $n$.
%
To summarize, we have
\begin{theorem}
  At the onset and bifurcation points,
  the square-free cyclic polynomials $C_p(\vx{n})$
  of the cycle points $\vx{n} = \{x_k\}$
  are linearly related as
  $\lambda \, C_p(\vx{n}) = \sum_{q \in \Pset} T_{pq}(R) \, C_q(\vx{n})$,
  with $\lambda$ being $+1$ and $-1$, respectively,
  and $T_{pq}(R)$ being the coefficients
  from the square-free reduction of the cyclic polynomial
  $C_p(\vx{n}) \, \prod_k f'(x_k)$. %[\refeq{logder}].
  Thus, the fixed points of $f^n$
  are marginally stable when $R$ satisfies
  the characteristic equation
  $\big| \vct T(R) - \lambda \, \vct I \big| = 0$.
  \label{thm:main}
\end{theorem}

@@

We illustrate the above algorithm on several small-$n$ cases.
%
%It is still helpful to have a mathematical software
%  at hand to verify certain steps
%  (e.g., in computing the determinants and their factorization).
%
%
For $n = 1$, % (the fixed point),
we have two square-free cyclic polynomials
$\{C_0, C_1\} = \{1, x_1\}$.
$\Lambda = -2 \, x_1$, and
\[
  \Lambda
  \left( \begin{array}{c}
  C_0 \\
  C_1
  \end{array} \right)
  =
  \left( \begin{array}{cc}
  0     & -2 \\
  -2R   &  2
  \end{array}\right)
  \left( \begin{array}{c}
  C_0 \\
  C_1
  \end{array} \right).
\]
The characteristic equation (\ref{eq:xr}) reads
\begin{equation}
A_1\Rlam
  %=\left| \begin{array}{cc}
  %\lambda     & 2          \\
  %2 R         & \lambda - 2
  %\end{array} \right|
  \equiv -4R - 2\lambda + \lambda^2
  = 0.
  \tag{\ref{eq:xr}-1}
\label{eq:xr1}
\end{equation}
So $R = (\lambda^2 -2\lambda)/4$,
  and the fixed point
  %begins at $R_a = -1/4$ ($\lambda = +1$), and
  becomes unstable at $R_b = 3/4$ where $\lambda = -1$.
